{\cis}: A Python Package for Integrating Computational Models Across Languages and Scales

### 2018-10-24 papers/lang2018/ms.tex
Add reference to FLAME

%---------------------------------------------------------------------------------
% DOMAIN SPECIFIC LANGUAGE
\subsection{Domain Specific Language (DSL)}\label{SS:dsl}
%
One solution is to agree upon a language. The questions of orchestration, communication, and translation become trivial when models are all written in the same language. Beyond selecting a single programming language, many communities have  developed dedicated domain specific languages (DSLs) for model representation. For example, the Systems Biology Markup Language \citep[SBML][]{Hucka2003} is an XML-based format for representing models of biological processes. Scientists can compose their models using SBML markup and then run their model using software designed to parse SBML files. Other DSLs for biological models include LPy \citep{Boudon2012}, CellML \citep{Cuellar2003}, and BioPAX \citep{Demir2010}. Dedicated model DSLs improve the reusability of models written in the DSL and have the advantage of often implicitly handling data transformation. However, they provide no help for existing models that are not in the DSL or those that cannot be expressed within the constraints of the DSL. In addition, learning a new DSL, in addition to a programming language, can be daunting.

>>>>

%---------------------------------------------------------------------------------
% DOMAIN SPECIFIC LANGUAGE
\subsection{Domain Specific Language (DSL)}\label{SS:dsl}
%
One solution is to agree upon a language. The questions of orchestration, communication, and translation become trivial when models are all written in the same language. Beyond selecting a single programming language, many communities have  developed dedicated domain specific languages (DSLs) for model representation. For example, the Systems Biology Markup Language \citep[SBML][]{Hucka2003} is an XML-based format for representing models of biological processes. Scientists can compose their models using SBML markup and then run their model using software designed to parse SBML files. Other DSLs for biological models include LPy \citep{Boudon2012}, CellML \citep{Cuellar2003}, FLAME \citep{Coakley2012}, and BioPAX \citep{Demir2010}. Dedicated model DSLs improve the reusability of models written in the DSL and have the advantage of often implicitly handling data transformation. However, they provide no help for existing models that are not written in the DSL or those that cannot be expressed within the constraints of the DSL. Furthermore, learning a new DSL, in addition to a programming language, can be daunting.
